Sketching for Density Estimation

### 2018-12-05 density.tex
some changes

\newcommand{\ip}[1]{\left \langle #1 \right \rangle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\rho}
\newcommand{\eps}{\epsilon}
%\usepackage{ntheorem}
\newcommand{\floor}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\ceil}[1]{\left \lceil #1 \right \rceil}

@@

For bounded functions $f$, uniform sampling combined with a union bound over $|\mathcal Q|$ always provides a valid solution. 
While $|\mathcal Q|$ is often infinite it reduces to a finite (albeit usually exponentially large) set through an epsilon net argument. 
This is common practice in machine learning, specifically, PAC learning. 
However, it is often far from being the best possible in terms of its space-accuracy tradeoff.

@@

We argue that we can offer a solution for producing coresets for all of the above problems in a unified manner. 
Moreover, our solution creates streaming algorithms with fully mergeable sketches.

The high level idea is the following. Consider a stream of the data points $x_i$. We wish to to maintain a buffer that can hold up to $k$ points, and output a stream of data points having half the amount of the original stream, yet at the same time maintain a similar value for 
$$ \sum_i f(x_i, q) $$
for any query $q$. To do so our buffer, once collecting $k$ data points, performs a \emph{compact} operation that outputs $k/2$ data points of weight $2$ representing the original $k$ inputs. The compact operation partitions the $k$ points into two (roughly) sets, whose corresponding sum are very close. This is made formal with the following definitions: Let us consider a signed sum of error function $E(q) = \sum_{i=1}^{n} s_i f(x_i,q)$ where $s_i \in \{-1,1\}$ are signs.
Now consider $F_1(q) = F(q) + E(q) = \sum_{i=1}^{n} f(x_i,q)  + \sum_{i=1}^{n} s_i f(x_i,q)  = 2 \sum_{i | s_i=1} f(x_i,q)$ and the corresponding $F_{-1}$. We have that both $F_1(q), F_{-1}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|F_1(q) - F(q)| \le  | E(q)|$.

@@

\subsection{optimal design of experiments}
We discuss the problem of linear regression with mean squared error. In the classic setting we are given $n$ points $x_1,\ldots,x_n \in \R^d$ along with $n$ labels $y_1,\ldots,y_n \in \R$ and wish to find $w \in \R^d$ minimizing
$$ \sum_i (w^Tx_i - y_i)^2 $$
In the field of optimal design of experiments, the labels are not given with the data. Instead, the learner must choose in advance a subset of datapoints for which they query the labels. The subset is chosen to provide the least error in a generalization bound, assuming $(x_1,y_1),\ldots,(x_n,y_n)$ are i.i.d from an unknown distribution.

@@

>>>>

\newcommand{\ip}[1]{\left \langle #1 \right \rangle}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\D}{\rho}
\newcommand{\eps}{\epsilon}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
%\usepackage{ntheorem}
\newcommand{\floor}[1]{\left \lfloor #1 \right \rfloor}
\newcommand{\ceil}[1]{\left \lceil #1 \right \rceil}

@@

Generating a concise representation for $\tilde F$ allows one to optimize of $\tilde F$ instead of $F$ which more efficient. 
Moreover, if the sketches or coresets are mergable, this could be done on separate machines without the need for communication or assuming randomness in the partitioning.

@@

For bounded functions $f$ like above, uniform sampling of $\ell = O(\log(1/\delta)/\eps^2)$ combined with a union bound over $|\mathcal Q|$ always provides a valid solution using $\ell = O(\log(|\mathcal Q|)/\eps^2)$ stream items. 
While $|\mathcal Q|$ is often infinite it reduces to a finite (albeit usually exponentially large) set through standard epsilon net arguments. 
This is common practice in machine learning, specifically, PAC learning. 
However, it is often far from being the smallest cardinality coreset possible.

We argue that we can offer a solution for producing smaller coresets for all of the above problems in a unified manner. 
Moreover, our solution creates streaming algorithms with fully mergeable sketches.

The high level idea is the following. 
Consider a stream of the data points $x_i$. 
We maintain a buffer that can hold up to $k$ points.
Once collecting $k$ data points it performs a \emph{compact} operation that outputs $k/2$ data points of weight $2$ representing the original $k$ inputs. The compact operation partitions the $k$ points into two sets whose corresponding sum are very close. 
This is made formal with the following definitions: 
Let us consider a signed sum of error function $E(q) = \sum_{i=1}^{n} s_i f(x_i,q)$ where $s_i \in \{-1,1\}$ are signs.
Now consider $F_1(q) = F(q) + E(q) = \sum_{i=1}^{n} f(x_i,q)  + \sum_{i=1}^{n} s_i f(x_i,q)  = 2 \sum_{i | s_i=1} f(x_i,q)$ and the corresponding $F_{-1}$. We have that both $F_1(q), F_{-1}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|F_1(q) - F(q)| \le  | E(q)|$.

@@

\subsection{Optimal Design of Experiments}
We discuss the problem of linear regression with mean squared error. In the classic setting we are given $n$ points $x_1,\ldots,x_n \in \R^d$ along with $n$ labels $y_1,\ldots,y_n \in \R$ and wish to find $w \in \R^d$ minimizing
$$ \sum_i (w^Tx_i - y_i)^2 $$
In the field of optimal design of experiments, the labels are not given with the data. Instead, the learner must choose in advance a subset of datapoints for which they query the labels. The subset is chosen to provide the least error in a generalization bound, assuming $(x_1,y_1),\ldots,(x_n,y_n)$ are i.i.d from an unknown distribution.

@@

\section{Attempt at Cleaner Definitions}
Let $\mathcal F$ be a family functions from a domain $\X$ to $\R$.
We define the discrepancy of $\F$. Let $f \subset{\F}$ to a subset of $\F$.
$$\mathcal D_m(\F) := \sup_{x\in\X} \min_{s\in \{0,1\}^m} \sup_{f\subset \F, |f|=m}  \sum_{i}s_i f_i (x)$$  \el{need to double check the order here...}

We claim that a very common task is to approximate the function $F = \sum_i f_i$ using a concise representation.  
Specifically, using coresets we can approximate $F$ with $\tilde F = \sum_i w_i f_i$ where $w$ is non-zero in at most $k$ places.

In this paper we show that if a family of functions exhibits low discrepancy then finding a good coreset is always possible.
The size of the coresets we find with the proposed algorithm is always smaller than that achievable by uniform sampling.
Moreover, we show that this framework includes a wide variety of problems including quantile approximation, approximate counting, machine learning classification and regressing, kernel density estimation and many more. 
Finally, we show that coresets can always be created in a streaming fashion using fully mergable sketches with only a $\operatorname{poly}\log\log$ factor loss in space.

In what follows, we subscript the function $f$ with its parametrization.
\begin{itemize}
\item In approximate counting we have $f_a(x) = 1$ if $a=x$ and zero else. Here both $a$ and $x$ belong to some finite domain.
\item In quantile approximation we have $f_a(x) = 1$ if $a<x$ and zero else. Here, $a$ and $x$ belong to a set which exhibits total ordering.
\item In matrix column subset selection we have $f_z(a,q) = \langle z,x \rangle ^2$ where $a,x \R^d$. For simplicity, we assume $\|a\| = \|x\| = 1$.
\item In density estimation $f_a(x) = \exp(- \|a-x\|_2^2/\sigma^2)$ or any other kernel (see \cite{})
\item In logistic regression $f_a(x) = \operatorname{loss}(a,x) = \log(1 + \exp(-a^T x))$. 
Here $x$ is the parameters of the linear model and $a$ is a training example vector multiplied by its label ($1$ or $-1$).
\item Linear Classification $f_a(x) = 1$ if $a^Tx > 0$ and $0$ else. Here, like in Logistic Regression, $a$ is the example point multiplied by its label and $x$ is the weights of the linear classifier.
\item Fractional Satisfiability $f(x,q)$ if $q$ satisfies $f$ and $0$ otherwise. Here $x$ in a clause in conjunctive normal form and $q$ is a boolean value assignment to the variables. For example $f(x,q) = 1$ iff $(q_1 \vee q_5  \vee \neg q_8)$ is true.
\item Quadratic forms of the graph Laplacian are $f(x, q) = q^T L_x q = \sum_{i,j\in E} (q_i - q_i)^2$ where $x$ is an edge and $L_x$ is the graph laplacian corresponding to the graph with a single edge $x$ and $q$ is a test vector.
\item Probably also good for matrix approximation when we get entry updates. 
\end{itemize}
